MG7: Configurable and scalable 16S metagenomics data analysis -- new methods optimized for massive cloud computing

### 2015-09-13 docs/paper/out/paper.tex
add new version of 16S db from @rtobes -- paper
See https://github.com/ohnosequences/metagenomica/issues/26#issuecomment-139846932

Our 16S Reference Database is a curated subset of sequences from NCBI
nucleotide database \textbf{nt}. This subset of 16S sequences was
selected by similarity with the bacterial and archaeal reference
sequences downloaded from RDP database \citep{cole2013ribosomal}. RDP
unaligned sequences were used to capture new16S sequences from nt using
BLAST similarity strategies and, then, performing additional curation
steps to remove sequences with poor taxonomic assignments to taxonomic
nodes close to the root of the taxonomic tree. All the nucleotide
sequences included in nt database has a taxonomic assignment provided by
the genbank sequence submitter. NCBI provides a table (available at
ftp://ftp.ncbi.nlm.nih.gov/pub/taxonomy/) to do the mapping of any
Genbank Identifier (GI) to its Taxonomy Identifier (TaxID). Thus, we are
based on a submitter-maintained taxonomic annotation system for
reference sequences that supposes a sustainable system able to face the
expected number of reference sequences that will populate the public
global nucleotide databases in the near future. Another advantageous
point is that we are based on NCBI taxonomy, the de facto standard
taxonomic classification for biomolecular data \citep{cochrane20102010}.
NCBI taxonomy is, without any doubt, the most used taxonomy all over the
world and the most similar to the official taxonomies of each specific
field. This is a crucial point because all the type-culture and tissue
databanks follow this official taxonomical classification and, in
addition, all the knowledge accumulated is referred to this taxonomy. In
addition NCBI provides a direct connection between taxonomical formal
names and the physical specimens that serve as exemplars for the species
\citep{federhen2014type}.

If metagenomics results are easily integrated with the theoretical and
experimental knowledge of each specific area, the impact of metagenomics
will be higher that if metagenomics progress in a disconnected research
branch. This strategy for building our database allows substituting the
16S database by any other subset of nt, even by the complete nt database
if it would needed, for example, for analyzing shotgun metagenomics
data.

>>>>

Our 16S Reference Database is a curated subset of sequences from NCBI
nucleotide database \textbf{nt}. The sequences included were selected by
similarity with the bacterial and archaeal reference sequences
downloaded from the \textbf{RDP database} \citep{cole2013ribosomal}. RDP
unaligned sequences were used to capture new 16S sequences from
\textbf{nt} using BLAST similarity search strategies and then,
performing additional curation steps to remove sequences with poor
taxonomic assignments to taxonomic nodes close to the root of the
taxonomy tree. All the nucleotide sequences included in \textbf{nt}
database has a taxonomic assignment provided by the \textbf{Genbank}
sequence submitter. NCBI provides a table (available at
ftp://ftp.ncbi.nlm.nih.gov/pub/taxonomy/) to do the mapping of any
Genbank Identifier (GI) to its Taxonomy Identifier (TaxID). Thus, we are
based on a crowdsourced submitter-maintained taxonomic annotation system
for reference sequences. It supposes a sustainable system able to face
the expected number of reference sequences that will populate the public
global nucleotide databases in the near future. Another advantageous
point is that we are based on NCBI taxonomy, the \emph{de facto}
standard taxonomic classification for biomolecular data
\citep{cochrane20102010}. NCBI taxonomy is, undoubtedly, the most used
taxonomy all over the world and the most similar to the official
taxonomies of each specific field. This is a crucial point because all
the type-culture and tissue databanks follow this official taxonomical
classification and, in addition, all the knowledge accumulated during
last decades is referred to this taxonomy. In addition NCBI provides a
direct connection between taxonomical formal names and the physical
specimens that serve as exemplars for the species
\citep{federhen2014type}.

Certainly, if metagenomics results are easily integrated with the
theoretical and experimental knowledge of each specific area, the impact
of metagenomics will be higher that if metagenomics progresses as a
disconnected research branch. Considering that metagenomics data
interoperability, which is especially critical in clinical environments,
requires a stable taxonomy to be used as reference, we decided to rely
on the most widely used taxonomy: the NCBI taxonomy. In addition, the
biggest global sequence database GenBank follows this taxonomy to
register the origin of all their submitted sequences. Our 16S database
building strategy allows the substitution of the 16S database by any
other subset of \textbf{nt}, even by the complete \textbf{nt} database
if it would be needed, for example, for analyzing shotgun metagenomics
data. This possibility of changing the reference database provides
flexibility to the system enabling it for easy updating and
project-driven personalization.
